Extracting sunbeams from cucumbers: Computational reproducibility is essential for archaeological science

### 2025-04-16 analysis/paper/paper.qmd
update with data source

The entire R code [@rcoreteamLanguageEnvironmentStatistical2024] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/xxx/xxx/xxx to enable re-use of materials and improve reproducibility and transparency [@marwickComputationalReproducibilityArchaeological2017]. All the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this compendium [@doi:10.1080/00031305.2017.1375986]. The R code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.

@@

```{r}
# reporting on my AER work
jas_rr_data <- 
  read_csv(here::here("analysis/data/JAS AER data analysis.csv"))

@@

I have presented a bibliometric analysis on the status of archaeology as a science, showing distinct disunity that is increasing over time. On average we generally behave as social scientists, with some elements in common with harder sciences. These observations are consistent with @Lakatos1978model's of a research program as a central foundation of irrefutable core assumptions complemented by a set of hypotheses, models, and methods that are adjusted, modified, or replaced by day-to-day research. Archaeology consists of multiple programs like this, with distinct and sometimes non-overlapping sets of core assumptions, such as processual and post-processual archaeology. Some programs are more amenable to reproducibility, while others offer insights through alternative methods. Among the programs that depend on quantitative methods to assess hypotheses and models, if they are to continue to progress through increased consensus through the accumulation of reliable facts and methods, it is essential for researchers to take computational reproducibility seriously. Computers have become a central field and laboratory instrument for much of our work, so we have an ethical duty to document how we change our data as it flows through silicon just as carefully as we document the operating parameters of a mass spectrometer or any other field or laboratory instrument. However, the current state of quantitative archaeology, with most researchers not using open source code, is comparable to the secrecy of alchemy prior to the emergence of chemistry. Abandoning this habit of secrecy in favour of transparency and reproducibility is vital if we are to avoid a future where our journals are filled with pretty pictures that the reader has no hope of repeating or adapting in their own work. Computational reproducibility must be considered a minimum requirement for evaluating the integrity and usefulness of empirical results.

@@

The data that support the findings of this study are openly available in [repository name] at http://doi.org/[doi]

>>>>

The entire R code [@rcoreteamLanguageEnvironmentStatistical2024] used for all the analysis and visualizations contained in this paper is included in the Supplementary Online Materials at https://doi.org/10.5281/zenodo.14897253 to enable re-use of materials and improve reproducibility and transparency [@marwickComputationalReproducibilityArchaeological2017]. All the figures, tables, and statistical test results presented here can be independently reproduced with the code and data in this compendium [@doi:10.1080/00031305.2017.1375986]. The R code is released under the MIT license, the data as CC-0, and figures as CC-BY, to enable maximum re-use.

@@

```{r}
# reporting on my AER work
# initially collected here: 
# https://docs.google.com/spreadsheets/d/1o4jSZ__OoCDssf0lWWeBheNEbgiheU_1Udj5X82VOvY/edit?gid=0#gid=0
jas_rr_data <- 
  read_csv(here::here("analysis/data/JAS AER data analysis.csv"))

@@

I have presented a bibliometric analysis on the status of archaeology as a science, showing distinct disunity that is increasing over time. On average we generally behave as social scientists, with some elements in common with harder sciences. These observations are consistent with @Lakatos1978's model of a research program as a central foundation of irrefutable core assumptions complemented by a set of hypotheses, models, and methods that are adjusted, modified, or replaced by day-to-day research. Archaeology consists of multiple programs like this, with distinct and sometimes non-overlapping sets of core assumptions, such as processual and post-processual archaeology. Some programs are more amenable to reproducibility, while others offer insights through alternative methods. Among the programs that depend on quantitative methods to assess hypotheses and models, if they are to continue to progress through increased consensus through the accumulation of reliable facts and methods, it is essential for researchers to take computational reproducibility seriously. Computers have become a central field and laboratory instrument for much of our work, so we have an ethical duty to document how we change our data as it flows through silicon just as carefully as we document the operating parameters of a mass spectrometer or any other field or laboratory instrument. However, the current state of quantitative archaeology, with most researchers not using open source code, is comparable to the secrecy of alchemy prior to the emergence of chemistry. Abandoning this habit of secrecy in favour of transparency and reproducibility is vital if we are to avoid a future where our journals are filled with pretty pictures that the reader has no hope of repeating or adapting in their own work. Computational reproducibility must be considered a minimum requirement for evaluating the integrity and usefulness of empirical results.

@@

The data that support the findings of this study are openly available in Zenodo at http://doi.org/10.5281/zenodo.14897253
